---
title: "Boxplot Solutions"
author: 
date: January 9, 2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

For help, see `?boxplot`, `?geom_boxplot` and/or

https://edav.info/box.html


Datasets:

`movies` -- `ggplot2movies` package: 58788 rows, 24 variables

```{r}
# uncomment and run once:
# install.packages("ggplot2movies")
```

Data: `countries2012.csv` 

Available in the data folder, direct link:

https://github.com/DS-BootCamp-DSI-Columbia/AY2018-2019-Winter-Collaboratory-Faculty/blob/master/Bootcamp-materials/Day1-EDAV-ggplot2/data/countries2012.csv

### 1. Single boxplots

Do any of the following:

a) Draw a boxplot of the `TFR` (total fertility rate) column in the countries dataset.

```{r}
library(tidyverse)
df <- read_csv("../data/countries2012.csv")
boxplot(df$TFR)
```

b) Draw a horizontal boxplot of the `length` variable of the `movies` dataset.

```{r}
library(ggplot2movies)
boxplot(movies$length, horizontal = TRUE)
```

Boxplot of movies under 3 hours:

```{r}
library(tidyverse)
movies2 <- movies %>% filter(length < 180)
boxplot(movies2$length, horizontal = TRUE)
```

(Not very informative...)

### 2. Multiple boxplots

a) Draw multiple boxplots of `TFR` in the countries dataset by `CONTINENT`.

```{r}
ggplot(df, aes(x = CONTINENT,     
                  y = TFR)) + 
  geom_boxplot()
```

Ordered by median in decreasing order:

```{r}
ggplot(df, aes(x = fct_reorder(CONTINENT, -TFR, median),
               y = TFR)) +
  geom_boxplot()
               
```


b) Draw multiple boxplots of `length` in the `movies` dataset by `mpaa`.

```{r}
ggplot(movies, aes(mpaa, length)) + 
  geom_boxplot()
```


c) Filter / subset the data in b) to obtain clearer views of the trends.

```{r}
# remove movies that don't have an MPAA rating
movies2 <- movies %>% filter(mpaa != "")

# use logical order for MPAA levels (move NC-17 to th end)

movies2 <- movies2 %>% 
  mutate(mpaa = fct_relevel(mpaa, "NC-17", after = Inf))

ggplot(movies2, aes(mpaa, length)) + geom_boxplot()
```

