---
title: "Barcharts"
author: 
date: January 9, 2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```


### 1. Recoding

Data:  `flower` dataset in  **cluster** package

Rename the column names and recode the levels of categorical variables to descriptive names. For example, "V1" should be renamed "winters" and the levels to "no" or "yes". Display the full dataset.

Help: `?fct_recode`

```{r}
library(cluster)
library(tidyverse)
f <- flower %>% 
  transmute(winter = fct_recode(V1, no = "0", yes = "1"),
  shadow = fct_recode(V2, no = "0", yes = "1"),
  tubers = fct_recode(V3, no = "0", yes = "1"),
  color = fct_recode(V4, white = "1", yellow = "2", pink = "3", red = "4", blue = "5"),
  soil = fct_recode(V5, dry = "1", normal = "2", wet = "3"),
  ranking = V6,
  height = V7,
  distance = V8)
f
```

### 2. Barcharts

Create frequency bar charts for the `color` and `soil` variables, using best practices for the order of the bars.

Help:

https://github.com/jtr13/codehelp/blob/master/R/reorder.md

```{r}
ggplot(f, aes(fct_infreq(color))) + geom_bar() + theme_grey(14)
```

```{r}
ggplot(f, aes(soil)) + 
  geom_bar() +
  scale_y_continuous(breaks = seq(0,10,2)) + theme_grey(14)
```

