---
title: "Continuous Variables"
output: slidy_presentation
---
```{r setup, include=FALSE, cache=FALSE}
library(knitr)
library(tidyverse)
options(digits = 3, scipen = 999)
opts_chunk$set(fig.align='center', 
               cache=TRUE, message = FALSE, 
               echo = FALSE)
opts_knit$set(global.par = TRUE)
```

## Continuous Variables
We're looking for features such as:

> - Asymmetry

> - Outliers

> - Multimodality

> - Gaps

> - Heaping / Rounding

> - Impossibilities / Errors



## Histograms

- primary tool for continuous data

> - count / relative frequency / density

> - boundaries

> - binwidth

## How are histograms created?

Draw a histogram on paper of the following data.  

(use binwidth = 5)

<h2>50, 51, 53, 55, 56, 60, 65, 65, 68</h2>

## How are histograms created?

<table><tr><td>
```{r, fig.width = 4, fig.height = 3, echo = FALSE}
par(las = 1)  # opts_knit$set(global.par = TRUE) above
x <- c(50, 51, 53, 55, 56, 60, 65, 65, 68)
hist(x, col = "lightblue")
```
</td><td>
```{r, fig.width = 4, fig.height = 3, echo = FALSE}
df <- data.frame(x)
ggplot(df, aes(x)) + geom_histogram(color = "black", fill = "lightblue", binwidth = 5, center = 52.5)
```
</td></tr></table>


## How are histograms created?

```{r, fig.align = 'left'}
x <- c(50, 51, 53, 55, 56, 60, 65, 65, 68)
hist(x, col = "lightblue", right = FALSE, yaxt = "n")
axis(2, at = 0:3)
```

## What is causing the difference?
```{r, fig.width = 8}
op <- par(mfrow = c(1, 2), las = 1)
hist(x, col = "lightblue", ylim = c(0, 4))
hist(x, col = "lightblue", right = FALSE, ylim = c(0, 4))
par(op)
```

## Bin boundaries
```{r, fig.width = 8}
op <- par(mfrow = c(1, 2), las = 1)
hist(x, col = "lightblue", ylim = c(0, 4),
     xlab = "right closed ex. (55, 60]", font.lab = 2)
hist(x, col = "lightblue", right = FALSE, ylim = c(0, 4),
     xlab = "right open ex. [55, 60)", font.lab = 2)
par(op)
```

## Bin boundaries
```{r, fig.width = 6, fig.height = 4}
# presentation issues
hist(x, breaks = seq(47.5, 72.5, 5), col = "lightblue",
     axes = FALSE)
axis(1, at = seq(47.5, 72.5, 5))
axis(2, at = 0:3, las = 1)
```

## Frequency (count) histogram 

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665, 675, 699, 699, 725, 725, 745, 799)
x <- hist(prices, breaks = seq(300, 800, 100),
          col = "lightblue", las = 1)
options(digits = 5)
df <- data.frame(mids = x$mids, freq = x$counts, 
                 relfreq = x$counts/sum(x$counts), 
                 density = x$density) 
```

Prices of one-bedroom apartments in Morningside Heights (zip 10027) in $1000k

## Relative frequency histogram

```{r}
x <- barplot(df$relfreq, space = 0, col = "lightblue",
             xlab = "prices",
             ylab = "relative frequency", las = 1)
# axis(1) to see the scale on the x-axis
axis(1, at = 0:5, labels = seq(300, 800, 100))
title("Relative Frequency Histogram of Prices")
```

## Density Histogram
```{r}
hist(prices, breaks = c(300, 400, 500, 600, 700, 800),
     freq = FALSE, col = "lightblue", ylab = "",
     main = "Density Histogram", las = 1)
mtext("Density", side = 2, line = 4, las = 3)
```

## Count, Relative Frequency, Density
```{r freq, eval = FALSE, echo = FALSE}
kable(df)
```

|Bin|Count|Relative Frequency|Density|
|:-----:|:-----:|:-----:|:-----:|
|300-400|1|.056|.00056|
|400-500|4|.22|.0022|
|500-600|5|.28|.0028|
|600-700|4|.22|.0022|
|700-800|4|.22|.0022|

> - How is relative frequency calculated?

> - How is density calculated?



## What's wrong with this histogram?
```{r}
# Use geom_col since we already have frequency counts
# This is an example of what not to do
df <- read.csv("../data/zip10027census2000.csv")
df$age <- factor(df$age, levels = df$age)
g0 <- ggplot(df, aes(x = age, y = pop)) + 
    geom_col(width = 1, color = "blue", fill = "lightblue") +
    ylab("count")
g0
```

## Relative frequency histogram

(still wrong...)

```{r}
# Doesn't fix the problem
ggplot(df, aes(x = age, y = percent/100)) + 
    geom_col(width = 1, color = "blue", fill = "lightblue") +
    ylab("relative frequency")
```

## Density histogram with unequal bin (or class) widths
```{r}
g2 <- ggplot(df, aes(x = center, y = percent/(100*binwidth), 
               width = binwidth)) + 
    geom_col(color = "blue", fill = "lightblue") +
    ylab("density") + xlab("age") +
    scale_x_continuous(breaks = c(0, df$breaks)) +
    ggtitle("Census 2000: Zip Code 10027")
g2
```

## Density = RelFreq / Binwidth
```{r}
library(dplyr)
kdf <- df %>% transmute(Class = age, Frequency = pop,
                        RelFreq = round(pop/sum(pop),3),
                        ClassWidth = binwidth,
                        Density = round(RelFreq/ClassWidth,3))
knitr::kable(kdf)
```

## Compare the histograms
```{r, fig.height = 4}
library(gridExtra)
grid.arrange(g0 + ggtitle ("Incorrect"),
             g2 + ggtitle ("Correct"))
```

Source:  https://factfinder.census.gov/

## Binwidth

<h2>`'stat_bin()' using 'bins = 30'. Pick better`
`value with 'binwidth'.`</h2>

## Histograms
```{r}
world <- read.csv("../data/countries2012.csv")
hist(world$TFR, col = "lightblue",
     main = "Total Fertility Rate", las = 1,
     xlab = "average total births per woman, by country")
```

## Histograms
```{r}
hist(world$TFR, col = "lightblue",
     main = "Total Fertility Rate", las = 1,
     breaks = 10,
     xlab = "average total births per woman, by country")
```


## Histograms
```{r}
hist(world$TFR, col = "lightblue",
     main = "Total Fertility Rate", las = 1,
     xlab = "average total births per woman, by country",
     breaks = seq(1 , 8, .25))
```

## Histograms
```{r}
hist(world$TFR, col = "lightblue",
     main = "Total Fertility Rate", las = 1,
     xlab = "average total births per woman, by country",
     breaks = seq(1 , 8, .125))
```

## Histograms
```{r}
df <- data.frame(scores = c(35, 59, 61, 64, 66, 66, 70, 72, 73, 74,
                            75, 76, 76, 78, 79, 80, 80, 81, 81, 82, 
                            82, 82, 84, 86, 86, 88, 88, 88, 88, 89, 
                            89, 90, 91,91, 92, 92, 92, 92, 94, 94, 
                            94, 94, 96, 98, 102))
ggplot(df, aes(x = scores)) + 
    geom_histogram(color = "blue", fill = "lightblue") +
    scale_x_continuous(breaks = seq(30, 100, 10)) +
    ggtitle("Test Scores")
```            
            
## Fewer bins
```{r}
ggplot(df, aes(x = scores)) + 
    geom_histogram(color = "blue", fill = "lightblue",
                   breaks = seq(30, 110, 10)) +
        scale_x_continuous(breaks = seq(30, 110, 10))
```


