---
title: "Histograms"
author: 
date: January 9, 2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

For help, see `?geom_histogram` and/or

https://edav.info/histo.html#parameters

### 1. Frequency histograms

Choose one of the following datasets:

`swiss` -- built-in base R: 47 rows, 6 variables

```{r}
str(swiss)
```

`movies` -- `ggplot2movies` package: 58788 rows, 24 variables

```{r}
# uncomment and run once:
# install.packages("ggplot2movies")
library(ggplot2movies)
str(movies)
```

Draw histograms of individual variables. Experiment with changing the bin width and center.

```{r}
library(ggplot2)

```

### 2. Add title and axis labels

Choose one graph and add a title and clear x and y axis labels (Use `?swiss` or `?movies` for more info about the datasets.)

### 3. Density histograms

Chooose one variable and create a density histogram with `geom_histogram(aes(y = ..density..))`


### 4. Density curve

Add a density curve with `+ geom_density()`

### 5. Normal curve

Challenge: Add a normal distribution curve. (Hint: see `?stat_function`).

### 6. Multiple density curves

Demo:

```{r}
swiss2 <- swiss %>% 
  mutate(edulevel = ifelse(Education < 9, 
                           "less than eight", 
                           "nine or more"))

ggplot(swiss2, aes(x = Fertility, 
                   color = edulevel, 
                   fill = edulevel)) +
  geom_density(alpha = .2) + 
  theme_classic()
```

Compare density curves for `length` of Action vs. non-Action movies. (Hint: convert the `Action` column to a factor)

