---
title: "Boxplots"
author: 
date: January 9, 2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

For help, see `?boxplot`, `?geom_boxplot` and/or

https://edav.info/box.html


Datasets:

`swiss` -- built-in base R: 47 rows, 6 variables

```{r}
str(swiss)
```

`movies` -- `ggplot2movies` package: 58788 rows, 24 variables

```{r}
# uncomment and run once:
# install.packages("ggplot2movies")
library(ggplot2movies)
str(movies)
```

### 1. Single boxplots

Do any of the following:

a) Draw a boxplot of the `Fertility` column in the `swiss` dataset.

b) Draw a horizontal boxplot of the `length` variable of the `movies` dataset.

### 2. Multiple boxplots

a) Draw multiple boxplots of `Fertility` in the `swiss` dataset by `Education` group.

b) Draw multiple boxplots of `length` in the `movies` dataset by `mpaa`.

c) Filter / subset the data in b) to obtain clearer views of the trends.
