---
title: "Boxplot"
author: 
date: January 9, 2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

For help, see `?boxplot`, `?geom_boxplot` and/or

https://edav.info/box.html


Datasets:

`movies` -- `ggplot2movies` package: 58788 rows, 24 variables

```{r}
# uncomment and run once:
# install.packages("ggplot2movies")
```

Data: `countries2012.csv` 

Available in the data folder, direct link:

https://github.com/DS-BootCamp-DSI-Columbia/AY2018-2019-Winter-Collaboratory-Faculty/blob/master/Bootcamp-materials/Day1-EDAV-ggplot2/data/countries2012.csv

### 1. Single boxplots

a) Draw a boxplot of the `TFR` (total fertility rate) column in the countries dataset.

b) Draw a horizontal boxplot of the `length` variable of the `movies` dataset.

### 2. Multiple boxplots

a) Draw multiple boxplots of `TFR` in the countries dataset by `CONTINENT`.

b) Draw multiple boxplots of `length` in the `movies` dataset by `mpaa`.

c) Filter / subset the data in b) to obtain clearer views of the trends.
