---
title: "Boxplot"
author: 
date: January 9, 2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

For help, see `?boxplot`, `?geom_boxplot` and/or

https://edav.info/box.html


Datasets:

`movies` -- `ggplot2movies` package: 58788 rows, 24 variables

```{r}
# uncomment and run once:
# install.packages("ggplot2movies")
```

Data: `countries2012.csv` 

Available in the data folder, direct link:

https://github.com/DS-BootCamp-DSI-Columbia/AY2018-2019-Winter-Collaboratory-Faculty/blob/master/Bootcamp-materials/Day1-EDAV-ggplot2/data/countries2012.csv

## Single boxplots

### 1. Countries

Draw a boxplot of the `TFR` (total fertility rate) column.

### 2. Movies

Draw a horizontal boxplot of the `length` variable.

## Multiple boxplots

### 3. Countries

Draw multiple boxplots of `TFR` by `CONTINENT`.

### 4. Movies

Draw multiple boxplots of `length` in the `movies` dataset by `mpaa`.

### 5. Filter / subset

Filter / subset the data in the previous question to obtain clearer views of the trends.
